---
title: "The Leslie Matrix - revisited"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

A population of fish we're interested in has four life stages: eggs (E), fry (F), juvenile (J), breeding adult (A). You are told that annually:

- Each breeding adult will produce, on average, 600 eggs
- 5% of eggs will survive to become fry
- 20% of fry will survive to become juveniles
- 16% of juveniles will survive to become adults
- Adult survival rate year to year is 90%

Recall the Leslie matrix from EDS 212. Matrix multiplication (using the dot product `%*%` of the matrix with the populations at each life stage) yields the population structure at the following year. 

### Create the Leslie matrix:

```{r}
# Make the Leslie Matrix:
fish_leslie <- matrix(c(0, 0, 0, 600, 0.05, 0, 0, 0, 0, 0.20, 0, 0, 0, 0, 0.16, 0.9), nrow = 4, ncol = 4, byrow = TRUE)

# Check it out: 
fish_leslie
```

The initial population structure is given by: 
```{r}
# Initial population structure (Roe, Fry, Juvenile, Adult): 
fish_year0 <- c(5000, 8000, 600, 400)
```


Add line-by-line comments wheres prompted in the code below, explaining what each is doing. Make sure that you are talking with your neighbors about what is really being created at each step. 

```{r}
# Model projections

# ------ WHAT HAPPENS HERE? ------ #
# ------ comment line-by-line ---- #

# The line below creates a time series to estimate pop size
time_yr <- seq(from = 0, to = 8, by = 1)

# The line below creates a matrix with dims that is the column length of fish_year0 and row length of time_year
proj_year <- matrix(ncol = length(fish_year0), nrow = length(time_yr))

# The line below puts the population of year zero in to the variable proj_year, the brackets allow us to index into a position. so it will store in the first row and has not specification for columns
proj_year[1, ] <- fish_year0

# The for loop below projects the population forward and stores them. Starts at two because first row is filled in and goes to the legnth of time_yr

#estimate pop sizer for each year based on pop size in the previous year
i<- 2
for (i in 2:length(time_yr)) { #start with year 1
  proj_year[i,] <- fish_leslie %*% proj_year[i-1,]
}


# The rest is wrangling & visualization (run to check it out):
#updating the names of the columns for the pop stages
colnames(proj_year) <- c("eggs", "fry", "juvenile", "adult")

# convert to a dataframe
proj_df <- data.frame(time_yr, proj_year) %>% 
  #convert format wide to long
  pivot_longer(cols = -time_yr, names_to = "lifestage", values_to = "stage_population")

#plot it
ggplot(data = proj_df, aes(x = time_yr, y = stage_population)) +
  geom_line(aes(color = lifestage)) +
  scale_y_log10()

```

Test different combinations of survival / reproduction and starting population structures to see what happens! 
