---
title: "Unit test practice"
format: html
editor: visual
---

# unit testing with the testthat package

First install necessary packages
```{r}
#install.packages("testthat")
#install.packages("devtools")
#install.packages("remotes")

library(testthat)
library(devtools)
library(remotes)
```

write a function to practice unit test

This function is going to find the mean value of each column, and then return the lowest and highest mean value calced (in that order)

We would expect:
- numeric
- output should be a vector legnth of two
-order should be lower and the other be higher

```{r}
mean_range <- function(df) {
  # will return the mean values of each column
  col_means <- apply(X = df, MARGIN = 2, FUN = mean, na.rm = TRUE)
  #combine min and max mean values
  col_mean_max <- max(col_means)
  col_mean_min <- min(col_means)
  
  #return the min and max in that order
  return(c(col_mean_min, col_mean_max))
} 

mean_range(df = mtcars)
```

Write some unit tests!
```{r}
#below is to hide the chunk output
#| include: false
#test the output is the length of 2
expect_length(mean_range(df = mtcars),2)
# nothing returned, meaning that its fine

#test that first value is less than second value

expect_true(mean_range(mtcars)[1] < mean_range(mtcars)[2])
```

## Practice sourcing functions
```{r}
#source our food_cart functions
#can only use r scripts
source(here::here("src","food_cart_functions.R"))

name_cart(food = "strawberry", animal = "marmot")
```




```{r}

```


